---
title: "PCA_AFKABAN"
author: "Muriel Dunn"
date: "2023-04-12"
output: html_document
---

```{r setup, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#tidyverse for cleaner code
library(tidyr)
library(tidyverse)
library(dplyr)
library(purrr)
library(feather)
library(vegan)
library(arrow)
library(fBasics)
```

## Clusters on broadband spectra
In this document I do a PCA to find the driving factor of the different clusters per species from AFKABAN experiments

## Prepped dataframes

Data from AFKABAN experiments are extracted into a dataframe in python (cluster_spectra.)

```{r data, warning=FALSE, message=FALSE}
a_cod120 = arrow::read_feather("./a_cod120pca.feather")
a_cod200 = arrow::read_feather("./a_cod200pca.feather")

p_cod120 = arrow::read_feather("./p_cod120pca.feather")
p_cod200 = arrow::read_feather("./p_cod200pca.feather")

p2_cod120 = arrow::read_feather("./p2_cod120pca.feather")
p2_cod200 = arrow::read_feather("./p2_cod200pca.feather")
```

## Run PCA

Write wrapper function to run PCA on normalize and centered data and plot results

```{r run_pca, warning=FALSE, message=FALSE}
######################################################################
#create a function to run pca and plot
run_pca <- function(data){
data_pca = princomp(data %>% select(2:last_col(1)), cor=T) # Cor standardizes data

summary(data_pca)

biplot(data_pca)

# Contruct biplot
colors = c("green", "blue", "red", "purple", "orange")
color_array = colors[data$y_kmean-min(data$y_kmean)+1]

plot(data_pca$scores[,1], data_pca$scores[,2],xlab="PCI",ylab="PCII", type="n") 
points(data_pca$scores[,1], data_pca$scores[,2],cex=1, col=color_array)# add color for different clusters
for (i in 1:length(data_pca$loadings[,1])){lines(c(0,data_pca$loadings[i,1]),c(0,data_pca$loadings[i,2]))}
text(data_pca$loadings[,1],data_pca$loadings[,2],labels=colnames(data %>% select(2:last_col(1))),cex=1.5,col="black")
return(data_pca)

}
##################################################################
```

```{r pca_acod, warning=FALSE, message=FALSE}
a_cod120_pca <- run_pca(a_cod120) 
summary(a_cod120_pca)
```

## Run for 200

```{r pca_200, warning=FALSE, message=FALSE}
a_cod200_pca <- run_pca(a_cod200) 
summary(a_cod200_pca)
```
## Run PCA on both polar cod experiments
```{r p_pca_120, warning=FALSE, message=FALSE}
p_cod120_pca <- run_pca(p_cod120) 
summary(p_cod120_pca)
```

```{r p_pca_200, warning=FALSE, message=FALSE}
p_cod200_pca <- run_pca(p_cod200) 
summary(p_cod200_pca)
```
```{r p2_pca_120, warning=FALSE, message=FALSE}
p2_cod120_pca <- run_pca(p2_cod120) 
summary(p2_cod120_pca)
```

```{r p2_pca_200, warning=FALSE, message=FALSE}
p2_cod200_pca <- run_pca(p2_cod200) 
summary(p2_cod200_pca)
``````