---
title: "Sound scattering models for: Model informed classification of broadband acoustic backscatter from zooplankton in an in situ mesocosm"
author: "Muriel Dunn"
date: '2023-11-16'
output: html_document
---


```{r setup, warning=FALSE, message=FALSE}
# tidyverse for cleaner code
library(tidyr)
library(dplyr)
library(purrr)
#apply with progress
library(pbapply)

#plotting
library(ggplot2)
library(gridExtra)

#special libraries
library(ZooScatR)


#pdf stuff
library(kableExtra)
library(knitr)

#exporting results
library(feather)

#This will run the simulations in parallel using with # cores available -1
runParallel = TRUE
if (runParallel == TRUE){
  library(foreach)
  library(doParallel)
  
  registerDoParallel(parallel::detectCores()-1)
}
#force running a new simulation, even if an RDS file with simulations is available
force=TRUE
#set working dir
setwd('C:\\Users\\mbd\\phd\\ZooScatStuff\\')
```
## AZKABAN population 

Copepods, euphausiids, chaetognaths and hydrozoans were the dominant species groups during the zooplankton mix experiment. For diverse Calanus groups we use literature values to populate the model parameters. For euphausiids, chaetognaths and hydrozoans we use the measurements from subset sampling for the length and width distributions.

## Model definition  

Loading the running parameters for the model:  

```{r setdir, warning=FALSE, message=FALSE}
#number of simulations
nsim = 1000
#location of the standard parameter file contained within the package
fname <- paste0(system.file(package="ZooScatR"),"/extdata/configs/config_0.dat")
# read the parameters file
para = ZooScatR::read_para(fname)
#set the soundspeed in the surrounding sea water
misc <- list(cw=1448.01)
#let's set the start and end frequencies

# Set the bandwidth
f0=45 #will bepara$simu$var0
f1=383 #will be para$simu$var1
# number of output frequencies
nf=2*(f1-f0)+1 #will be para$simu$n


```

For the parameter distributions a gamma, normal, lognormal or uniform distribution was used.  

The acoustically-dominant species that are considered weakly scatterers are the ones described in the table below. The table also contains all their parameters space settings. 

```{r model_set, warning=FALSE, message=FALSE}
pm = data.frame(species=c('Copepods','Euphausiids','Chaetognaths','Hydrozoans'),
  L_dist=c('rnorm', 'rlnorm','rgamma','rlnorm'),
  L_shape =  c(3.3, 2.4, 10.6, 2.4),
  L_rate =   c(0.7, 0.3,  0.6, 0.4),
  L_a_dist = c('rnorm'),
  L_a_mean = c(5.3, 11, 26, 2.8),
  L_a_sd =   c(0.9,  2, 8, 0.5),
  g_dist = 'rnorm',
  g_mean = c(0.997,1.037, 1.03, 1.007),
  g_sd   = c(0.005,0.005, 0.005, 0.005),
  h_dist = 'rnorm',
  h_mean = c(1.027,1.026, 1.03, 1.007),
  h_sd   = c(0.007,0.005, 0.005, 0.005),
  theta_dist = 'rnorm',
  theta_mean = c(90, 20,    0, 90),
  theta_sd =   c(30, 20,   30, 30),
  rho_l =    c(  3,   3,    0, 1),
  taper =    c(  2,   4,    4, 2) #taper =2 L/a>2 prolate spheroid, others selected from ZooScat Ex.
)
```
  
```{r table, echo=FALSE}
knitr::kable(pm[,1:7], booktabs = T, linesep = "") %>%
kable_styling(latex_options = "striped")

knitr::kable(pm[,c(1,8:13)], booktabs = T, linesep = "") %>%
kable_styling(latex_options = "striped")

knitr::kable(pm[,c(1,14:18)], booktabs = T, linesep = "") %>%
kable_styling(latex_options = "striped")

```

### Define the Model parameters

Now that we have lists of settings we can set those in the parameter files for each species group.  

#### Copepods (Calanus finmarchicus CV)

```{r get_para, warning=FALSE, message=FALSE}
######################################################################
#create a get_para funciton
get_para <- function(pm,spec,fn=-1){

pmsel = pm%>%filter(species==spec)
  if(pmsel$L_dist=='rnorm'){
          L = do.call(pmsel$L_dist,list(nsim,mean=pmsel$L_shape,sd=pmsel$L_rate))
    }else if (pmsel$L_dist=='rlnorm'){
          L = do.call(pmsel$L_dist,list(nsim,meanlog=pmsel$L_shape,sdlog=pmsel$L_rate))
    }else{
          L = do.call(pmsel$L_dist,list(nsim,shape=pmsel$L_shape,rate=pmsel$L_rate))
  }
  cp = data.frame(species=spec,
  L=L,
  La=do.call(pmsel$L_a_dist, list(nsim, pmsel$L_a_mean, pmsel$L_a_sd)),
  g=do.call(pmsel$g_dist,list(nsim, pmsel$g_mean, pmsel$g_sd)),
  h=do.call(pmsel$h_dist,list(nsim, pmsel$h_mean, pmsel$h_sd)),
  taper=pmsel$taper,
  rhol=pmsel$rho_l,
  pf=fn,
  theta=do.call(pmsel$theta_dist,list(nsim, pmsel$theta_mean, pmsel$theta_sd)))
  cdf = gather(cp,'var','value',-species,-taper, -rhol,-pf)
  subtit <- c('L' = "Length~(mm)",
  'La' = "L/a",
  'g' ="Density~contrast~g",
  'h' = 'Sound~speed~contrast~h',
  'theta' = 'Angle~theta~(degree)' )
  p <- ggplot(data=cdf, aes(x=value))+
  geom_density()+
  facet_wrap(.~var, scales='free')+#, labeller =labeller(var= as_labeller(subtit, label_parsed)))+
  theme_classic()+
  ggtitle(paste(spec,'- Material properties - Density plots'))+
  xlab('')+ylab('')

  return(list(cp,p))
}


##################################################################
#set copepod settings
fn = 'Shapes/Copepod_KB3_TT_150122_0134.sat'
cout = get_para(pm, 'Copepods', fn)
cdf = cout[[1]]
cout[[2]]
```

#### Krill (Thyssanoessa inermis)  

```{r krill, warning=FALSE, message=FALSE}
fn = 'Shapes/Euphausiid_KB3_TT_150122_0001-2.sat' # shape file from images
kpout = get_para(pm, "Euphausiids",fn)
kpdf = kpout[[1]]
kpout[[2]]
```

#### Chaetognath

#### Chaetognaths (Parasagitta )

```{r chaet, warning=FALSE, message=FALSE}
fn = 'Shapes/Chaetognath_KB3_TT_150122_0091.sat' # shape file from images
chout = get_para(pm, 'Chaetognaths',fn)
chdf = chout[[1]]
chout[[2]]
```
#### Jellyfish

#### Hydromedusae (Aglantha digitale)

```{r jelly, warning=FALSE, message=FALSE}
fn = 'Shapes/Hydrozoa_KB3_TT_150122_0026.sat'  # shape file from images
jfout = get_para(pm, 'Hydrozoan',fn)
jfdf = jfout[[1]]
jfout[[2]]
```


### Run the model

Now we can combine all the settings and prepare to run the simulations:

```{r combSettings, warning=FALSE, message=FALSE}
setdf = do.call('rbind',list(cdf,kpdf,chdf, jfdf))
set_para <- function(i, shape=FALSE, setdf=setdf){
  para = ZooScatR::read_para(fname)
  #let's set the start and end frequencies
  para$simu$var0=f0
  para$simu$var1=f1
  #number of output frequencies
  para$simu$n=nf
  para$shape$L = setdf$L[i]
  para$shape$L_a = setdf$La[i]
  para$phy$g0 = setdf$g[i]
  para$phy$h0 = setdf$h[i]
  para$shape$order = setdf$taper[i]
  para$shape$rho_L = setdf$rhol[i]
  para$shape$prof_name = setdf$pf[i]
  para$orient$angm = setdf$theta[i]
  
  para$shape$axis_sm = 80

  if(shape==FALSE){
    res = ZooScatR::bscat(para=para, misc=misc) #Target strength vs Frequency
    tmp=data.frame(TS=res$y,
    freq=seq(f0,f1, length.out = nf),
    L=para$shape$L,
    la=para$shape$L_a,
    g=para$phy$g0,
    h=para$phy$h0,
    orient=para$orient$angm,
    spec=setdf$species[i])
    return(tmp)
    
  }else{
    p = ZooScatR::buildpos(para)
    p = p$plot
    p = p+coord_equal()
    p = p+ggtitle(setdf$species[i])
    print(p)
    return(p)
  }
}

```

### Model input shapes
Plot the selected shapes:


```{r showShapes, fig.height=4,fig.width=8, warning=FALSE, message=FALSE}
# Shape arg means plot shape (y) or calculate backscatter using shape (n)
specp = setdf%>%group_by(species)%>%filter(row_number()==1)
for(i in 1:length(specp$species)){set_para(i,shape=TRUE, setdf=specp)}
```
```{r saveShapes, fig.height=4,fig.width=8, warning=FALSE, message=FALSE}
# Shape arg means plot shape (y) or calculate backscatter using shape (n)

c_shape = set_para(1,shape=TRUE, setdf=specp)
e_shape = set_para(2,shape=TRUE, setdf=specp)
ch_shape = set_para(3,shape=TRUE, setdf=specp)
h_shape = set_para(4,shape=TRUE, setdf=specp)

png(filename = 'shapes_ZoopMix.png')
grid.arrange(c_shape,e_shape,ch_shape,h_shape,nrow=2)
dev.off()
```

Run simulations:

```{r runSIm, warning=FALSE, message=FALSE}
if(force ==FALSE & file.exists('sim2.RDS')){
  message(Sys.time(), ': Found simulations...Using old simulaitons...')
  sim = readRDS('sim2.RDS')
}else{
message(Sys.time(),': Running simulations...')
if(runParallel == TRUE){
sim = foreach(i=1:length(setdf$L), .combine=rbind) %dopar% {set_para(i, shape=FALSE,setdf=setdf)}
}else{
sim = do.call('rbind',
pbapply(matrix(1:length(setdf$L)),1,function(i){set_para(i, shape=FALSE,setdf=setdf)})
)
}
message(Sys.time(),': Running simulations completed!')
saveRDS(sim, file='sim2.RDS', compress=TRUE)
}
```
  

Visualization
Plot Ts and length:
  
```{r fig.width=12, fig.height=12}
plot_func <- function(sim, name='L (mm)', var='L', y='TS',x='freq', title) {
  ggplot(data = sim, aes_string(x = x, y = y, col = var)) +
    geom_point(size=0.01) +
    scale_colour_viridis_c(name=name)+
    ggtitle(title)+
    xlab('Frequency (kHz)')+
    ylab('TS (dB)')+
    theme_classic()
}
pL <- function(sim,spec){plot_func(sim, name='L (mm)', var='L', y='TS',x='freq', spec)}

sim_tmp <- sim %>% 
  group_by(spec) %>% 
  nest() %>% 
  mutate(plots = map2(data, spec, pL))
pL = gridExtra::grid.arrange(grobs = sim_tmp$plots)
```


Density plots:  

```{r fig.width=12, fig.height=12}
pL <- function(sim, title){

  ggplot(data=sim, aes(x=freq, y=TS))+
    geom_density2d_filled()+
    #facet_wrap(.~spec, scales='free')+
    scale_x_continuous(expand=c(0,0))+
    scale_y_continuous(expand=c(0,0))+
    ggtitle(title)+
    theme_classic()
}

sim_tmp <- sim%>% 
  group_by(spec) %>% 
  nest() %>% 
  mutate(plots = map2(data, spec, pL))
pL = gridExtra::grid.arrange(grobs = sim_tmp$plots)
```

Export model simulations to feather file:

```{r export-feather}
path <- "AZKABAN_ZoopMix.feather"
write_feather(sim, path)

```
