---
title: "LengthDist_AZKABAN"
author: "Muriel Dunn"
date: '2022-05-19'
output: html_document
---

---
title: "Fit length ditributions_AZKABAN"
author: "Muriel Dunn"
date:  "`r Sys.Date()`"
output: html_document
---

```{r setup, warning=FALSE, message=FALSE}
library(fitdistrplus)
require(actuar)
require(gamlss.dist)
library(truncnorm)
library(feather)
```


## Fitting the distributions
I want to find the best fit for the length measurements I have from the multinet and tucker trawl data. A represenattive length distribution will inform and constrain the model input possibilities.

```{r values}

#Tucker Trawl @ KB3 Jan 15-16th Actual net values!
L_krill = c(7.14,7.14,8.73,8.73,8.73,8.73,8.73,8.73,8.73,8.73,8.73,8.73,8.73,8.73,8.73,8.73,8.73,9.52,9.52,9.52,9.52,9.52,9.52,9.52,9.52,9.52,9.52,9.52,9.52,9.52,10.32,10.32,10.32,10.32,10.32,10.32,10.32,10.32,10.32,10.32,11.11,11.11,11.11,11.11,11.11,11.11,11.11,11.11,11.11,11.11,11.11,11.11,11.11,11.11,12.70,12.70,12.70,12.70,12.70,12.70,12.70,12.70,12.70,12.70,13.49,13.49,13.49,14.29,14.29,17.46,19.05,19.05,21.43,22.22,23.81,25.40,26.19) #mm
# Krill from KB3 sample pictures
L_krill_pics = read_feather("L_Euphausiids.feather")

La_krill = read_feather("La_Euphausiids.feather")

# Jelly from KB3 preserved sample pictures + extra long lengths
L_jelly = read_feather("L_Hydrozoa.feather")
# Add measurements too large for pictures
L_jelly = c(L_jelly$L, 18, 17, 17, 26)
La_jelly = read_feather("La_Hydrozoa.feather")

# Chaetognaths from KB3 sample pictures + extra long lengths
L_chaet = read_feather("L_Chaetognaths.feather")
# Add measurements too large for pictures
L_chaet = c(L_chaet$L, 28,25,27,21,18,21,32,28,23,30,29,21,24,24,26,30,27,25,24,22)
La_chaet = read_feather("La_Chaetognaths.feather")

# Copepod from KB3 sample pictures
L_copepod = read_feather("L_Copepods.feather")

La_copepod = read_feather("La_Copepods.feather")
```



## Krill


```{r}
krill_dist = fitdist(c(L_krill), "lnorm")
length(L_krill)
mean(L_krill)
mean(L_krill_pics$L)
sd(L_krill)
sd(L_krill_pics$L)
plot(krill_dist)
summary(krill_dist)

# Pictures and measurements (MALIN) do not match but we use Malin's measurements -  more and representative of whole sample.
```
# Krill width

```{r}
krillla_dist = fitdist(c(La_krill$La), "norm")
length(La_krill$La)
mean(La_krill$La)
sd(La_krill$La)

plot(krillla_dist)
summary(krillla_dist)
```

## Hydrozoa
# Length
```{r}
jelly_dist = fitdist(L_jelly, "lnorm")
length(L_jelly)
mean(L_jelly)
sd(L_jelly)
plot(jelly_dist)
summary(jelly_dist)
```
# Fatness
```{r}
jellyla_dist = fitdist(La_jelly$La, "norm")
length(La_jelly$La)
mean(La_jelly$La)
sd(La_jelly$La)
plot(jellyla_dist)
summary(jellyla_dist)
```


## Chaetognaths
# Length
```{r}
chaet_dist = fitdist(L_chaet, "gamma")
length(L_chaet)
mean(L_chaet)
sd(L_chaet)
plot(chaet_dist)
summary(chaet_dist)
```
# Fatness
```{r}
chaetla_dist = fitdist(La_chaet$La, "norm")
length(La_chaet$La)
mean(La_chaet$La)
sd(La_chaet$La)
plot(chaetla_dist)
summary(chaetla_dist)
```



## Copepod
# Length
```{r}
copepod_dist = fitdist(L_copepod$L, "norm")
length(L_copepod$L)
mean(L_copepod$L)
sd(L_copepod$L)
plot(copepod_dist)
summary(copepod_dist)
```
# Fatness
```{r}
copepodla_dist = fitdist(La_copepod$La, "norm")
length(La_copepod$La)
mean(La_copepod$La)
sd(La_copepod$La)
plot(copepodla_dist)
summary(copepodla_dist)
```