---
title: "Running a DWBA simulation ensemble for Calanus finmarchicus in Tromsøflaket"
author: "Muriel Dunn"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
options(tibble.print_min = 4L, tibble.print_max = 4L)
library(ZooScatR)
library(dplyr)
library(reshape2)
library(ggplot2)
set.seed(1014)
```
This document shows how I used the DWBA model within the ZooScatR package to run model ensembles to calculate the target spectra of Calanus finmarchicus in Tromsøflaket.

## First iteration

I have entered the parameter files for an "average" Calanus finmarchicus in the config_calanusFinCV.dat. Let's check it out to see if it makes sense for TS spectra, backscattering cross-section and TS vs ka (to check for Rayleigh to geometric scattering.)
```{r fig.width=8,fig.height=4}
fname <- paste0("/Users/mbd/phd/ZooScatStuff/config_EukrohniaHamata.dat") #Location of the parameters file

para = read_para(fname) #Read parameters file
#Create list with soundspeed info
misc <- list(cw=1486)
#Run DWBA based on config file
res_TS <- bscat(para=para, misc=misc) #Target strength vs Frequency
res_TS$rplot #Show the result plot


#cross-section vs frequency
para$simu$out_indx <- 2 #Set output variable to cross section
#Run DWBA based on config file
res_sig <- bscat(para=para, misc=misc)
res_sig$rplot #Show the result plot

#Backscattering amplitude vs ka (Wavenumber x width)
para$simu$out_indx <- 3 #Set output to Target Strength
para$simu$var_indx <- 3 #Set output variable to ka
para$simu$var0 = 0.1 #simulate from ka=0 
para$simu$var1 = 10 #...to ka=10
#Run DWBA based on config file
res <- bscat(para=para, misc=misc)
res$rplot #Show the result plot
```


## SIMULATIONS!
```{r run_sim_auto}

Species_name = "CalanusFinCV"
Species_name = "EuphausiaceaFurcilia"
Species_name = "EukrohniaHamata"

fname <- paste0("/Users/mbd/phd/ZooScatStuff/config_", Species_name, ".dat")
para = read_para(fname) #Read parameters file

#make a length simulation
para$shape$ave_flag <- 1
#make an orientation simulation
para$orient$ave_flag <- 1

para$simu$var0 = 180 #simulate from 180 
para$simu$var1 = 450 #...to 450 kHz
para$simu$ni = 200 #reduce the number of elements and frequencies to improve speed
para$simu$n = 483

#Create list with sound speed info
misc <- list(cw=1486)



res <- bscat(para=para, misc=misc, simOut = TRUE, nang=350, nl=350) #Target strength vs Frequency
```


## Plotting the results
```{r plots_osim}
o_sim = melt(res$ysim)
o_sim$theta = res$ang[o_sim$Var2]
o_sim$Frequency = res$var[o_sim$Var1]
o_sim=o_sim[,3:ncol(o_sim)]
names(o_sim)[1] <- 'TS'

mTS = data.frame(Frequency=res$var,TS=res$y)

ggplot()+geom_line(data=o_sim, aes(x = Frequency, y=TS, group=theta),lty=2, lwd=0.5, alpha=0.4)+
  geom_line(data=mTS, aes(x=Frequency, y=TS), lty=1, lwd=2)+
  theme_classic()+
  theme(text=element_text(size=14))+
  ggtitle(Species_name)

ggplot(data=o_sim, aes(x=Frequency, y=theta, fill=TS))+
  geom_raster()+
  scale_fill_viridis_c()+
  scale_y_continuous(expand=c(0,0))+
  scale_x_continuous(expand=c(0,0))+
  theme_classic()+theme(text=element_text(size=14))+
  ggtitle(Species_name)
```  

### Length simulations

```{r plots_L}
l_sim =  melt(res$ysimL)
l_sim$Length = res$L[l_sim$Var2]
l_sim$Frequency = res$var[l_sim$Var1]
l_sim=l_sim[,3:ncol(l_sim)]
names(l_sim)[1] <- 'TS'

ggplot()+geom_line(data=l_sim, aes(x = Frequency, y=TS, group=Length),lty=2, lwd=0.5, alpha=0.4)+
  geom_line(data=mTS, aes(x=Frequency, y=TS), lty=1, lwd=2)+theme_classic()+theme(text=element_text(size=14))+
  ggtitle(Species_name)



ggplot(data=l_sim, aes(x=Frequency, y=Length, fill=TS))+
  geom_raster()+
  scale_fill_viridis_c()+
  scale_y_continuous(expand=c(0,0))+
  scale_x_continuous(expand=c(0,0))+
  theme_classic()+theme(text=element_text(size=14))+
  ggtitle(Species_name)

```



### Confidence Intervals of TS Spectra


```{r plots_CI}




```
