---
title: "ZoopMix_AZKABAN"
author: "Muriel Dunn"
date: '2022-05-19'
output: html_document
---


```{r setup, warning=FALSE, message=FALSE}
# tidyverse for cleaner code
library(tidyr)
library(dplyr)
library(purrr)
#apply with progress
library(pbapply)

#plotting
library(ggplot2)

#special libraries
library(ZooScatR)


#pdf stuff
library(kableExtra)
library(knitr)

#exporting results
library(feather)

#define if simulations should be multi-threadded. THis will run the simulations in parallel using with # cores available -1
runParallel = TRUE
if (runParallel == TRUE){
  library(foreach)
  library(doParallel)
  
  registerDoParallel(parallel::detectCores()-1)
}
#force running a new simulation, even if an RDS file with simualtions is available
force=TRUE
#set working dir
setwd('C:\\Users\\mbd\\phd\\ZooScatStuff\\')
```
## AZKABAN population 

Copepods, euphausiids, and chaetognaths were the dominant species groups during the zooplankton mix experiment. For diverse Calanus groups we can use literature values to populate the mode. For chaetognaths  and euphausiids we use the measurements from subset sampling.

## Model definition  

Let's start by loading some standard values for the model:  

```{r setdir, warning=FALSE, message=FALSE}
#number of simulations
nsim = 1000
#location of the standard parameter file contained within the package
fname <- paste0(system.file(package="ZooScatR"),"/extdata/configs/config_0.dat")
# read the parameters file
para = ZooScatR::read_para(fname)
#set the soundspeed in the surrounding sea water
misc <- list(cw=1446.56)
#let's set the start and end frequencies
f0=185 #will bepara$simu$var0
f1=255 #will be para$simu$var1
# number of output frequencies
nf=141 #will be para$simu$n
```

For the parameter distributions a gamma, normal or uniform dsitribution were used.  
For gamma distributions a density function with a shape $s$ and a rate $a$ was defined as:

$$f_{\Gamma}(x)= \frac{1}{s^a \Gamma(a) x^{a-1} e^{-\frac{x}{s}}}$$
where $\Gamma(a)$ is defined as [@milne-thomson_handbook_1972]:  
 $$\Gamma(a)=\int_0^\infty{t^{a-1}e^{-t} dt}$$

Normal distributions were computed with standard deviation $\sigma$ and mean $\mu$ with a density given by [@johnson_continuous_1995]:
$$f_N(x)=\frac{1}{\sqrt{2\pi}\sigma e^{-{\frac{(x-\mu)^2}{2\sigma^2}}}}$$
A summary of the settings is defined in table 1.
The acoustically-dominant species that are considered weakly scatterers are the ones described in the table below. the table also contains all their parameters space settings. 

```{r model_set, warning=FALSE, message=FALSE}
pm = data.frame(species=c('Copepods','Euphausiids','Chaetognaths','Jellyfish'),
  L_dist=c('rnorm', 'rlnorm','rlnorm','rnorm'),
  L_shape =  c(2.62, 2.4, 2, 2),
  L_rate =   c(0.09, 0.27,  0.4, 0.5),
  L_a_dist = c('rnorm'),
  L_a_mean = c(2.7, 10.5, 17.15, 2),
  L_a_sd =   c(0.2,  0.3, 0.5, 0.5),
  g_dist = 'rnorm',
  g_mean = c(0.996,1.037, 1.028, 1.014),
  g_sd = c(0.003,0.005, 0.005, 0.005),
  h_dist = 'rnorm',
  h_mean = c(1.027,1.026, 1.028, 1.014),
  h_sd = c(0.005,0.005, 0.005, 0.005),
  theta_dist = 'rnorm',
  theta_mean = c(90, 20,    0, 0),
  theta_sd =   c(30, 20,   30, 30),
  rho_l =    c(  3,   3,    3, 0),
  taper =    c(  2,   4,    2, 2) #taper =2 L/a>2 prolate spheroid, others selected from ZooScat Ex.
)
```
  
```{r table, echo=FALSE}
knitr::kable(pm[,1:7], booktabs = T, linesep = "") %>%
kable_styling(latex_options = "striped")

knitr::kable(pm[,c(1,8:13)], booktabs = T, linesep = "") %>%
kable_styling(latex_options = "striped")

knitr::kable(pm[,c(1,14:18)], booktabs = T, linesep = "") %>%
kable_styling(latex_options = "striped")

```

### Define the Model parameters

Now that we have lists of settings we can set those in the parameter files for each species group.  

#### Copepods (Calanus finmarchicus CV)

```{r get_para, warning=FALSE, message=FALSE}
######################################################################
#create a get_para funciton
get_para <- function(pm,spec,fn=-1){

pmsel = pm%>%filter(species==spec)
  if(pmsel$L_dist=='rnorm'){
          L = do.call(pmsel$L_dist,list(nsim,mean=pmsel$L_shape,sd=pmsel$L_rate))
    }else if (pmsel$L_dist=='rlnorm'){
          L = do.call(pmsel$L_dist,list(nsim,meanlog=pmsel$L_shape,sdlog=pmsel$L_rate))
    }else{
          L = do.call(pmsel$L_dist,list(nsim,shape=pmsel$L_shape,rate=pmsel$L_rate))
  }
  cp = data.frame(species=spec,
  L=L,
  La=do.call(pmsel$L_a_dist, list(nsim, pmsel$L_a_mean, pmsel$L_a_sd)),
  g=do.call(pmsel$g_dist,list(nsim, pmsel$g_mean, pmsel$g_sd)),
  h=do.call(pmsel$h_dist,list(nsim, pmsel$h_mean, pmsel$h_sd)),
  taper=pmsel$taper,
  rhol=pmsel$rho_l,
  pf=fn,
  theta=do.call(pmsel$theta_dist,list(nsim, pmsel$theta_mean, pmsel$theta_sd)))
  cdf = gather(cp,'var','value',-species,-taper, -rhol,-pf)
  subtit <- c('L' = "Length~(mm)",
  'La' = "L/a",
  'g' ="Density~contrast~g",
  'h' = 'Sound~speed~contrast~h',
  'theta' = 'Angle~theta~(degree)' )
  p <- ggplot(data=cdf, aes(x=value))+
  geom_density()+
  facet_wrap(.~var, scales='free')+#, labeller =labeller(var= as_labeller(subtit, label_parsed)))+
  theme_classic()+
  ggtitle(paste(spec,'- Material properties - Density plots'))+
  xlab('')+ylab('')

  return(list(cp,p))
}
##################################################################
#set copepod settings
#fn = paste0(system.file(package="ZooScatR"),"/extdata/profiles/copepod0.dat")
fn = 'cop0.sat'
cout = get_para(pm, 'Copepods', fn)
cdf = cout[[1]]
cout[[2]]
```


#### Krill (Thyssanoessa inermis)  

```{r krill, warning=FALSE, message=FALSE}
fn = paste0(system.file(package="ZooScatR"),"/extdata/profiles/euphaus0.dat")
kpout = get_para(pm, "Euphausiids",fn)
kpdf = kpout[[1]]
kpout[[2]]
```

#### Chaetognath

#### Chaetognaths (Parasagitta )

```{r chaet, warning=FALSE, message=FALSE}
fn = 'chaeto0.sat'
chout = get_para(pm, 'Chaetognaths',fn)
chdf = chout[[1]]
chout[[2]]
```
#### Jellyfish

#### Hydromedusae (Aglantha digitale)

```{r jelly, warning=FALSE, message=FALSE}
fn = 'Aglanthadigitale_shape_Moroz2014.sat'
jfout = get_para(pm, 'Jellyfish',fn)
jfdf = jfout[[1]]
jfout[[2]]
```


### Run the model

Now we can combine all the settings and prepare to run the simulations:

```{r combSettings, warning=FALSE, message=FALSE}
setdf = do.call('rbind',list(cdf,kpdf,chdf, jfdf))
set_para <- function(i, shape=FALSE, setdf=setdf){
  para = ZooScatR::read_para(fname)
  #set the soundspeed in the surrounding sea water
  #misc <- list(cw=1500)
  #let's set the start and end frequencies
  para$simu$var0=f0
  para$simu$var1=f1
  #number of output frequencies
  para$simu$n=nf
  para$shape$L = setdf$L[i]
  para$shape$L_a = setdf$La[i]
  para$phy$g0 = setdf$g[i]
  para$phy$h0 = setdf$h[i]
  para$shape$order = setdf$taper[i]
  para$shape$rho_L = setdf$rhol[i]
  para$shape$prof_name = setdf$pf[i]
  para$orient$angm = setdf$theta[i]
  
  if(shape==FALSE){
    res = ZooScatR::bscat(para=para, misc=misc) #Target strength vs Frequency
    tmp=data.frame(TS=res$y,
    freq=seq(f0,f1, length.out = nf),
    L=para$shape$L,
    la=para$shape$L_a,
    g=para$phy$g0,
    h=para$phy$h0,
    orient=para$orient$angm,
    spec=setdf$species[i])
    return(tmp)
    
  }else{
    p = ZooScatR::buildpos(para)
    p = p$plot
    p = p+coord_equal()
    p = p+ggtitle(setdf$species[i])
    print(p)
    return(p)
  }
}

```

### Model input shapes

First let’s have a look at the selected shapes:

```{r showShapes, fig.height=4,fig.width=8, warning=FALSE, message=FALSE}
specp = setdf%>%group_by(species)%>%filter(row_number()==1)
for(i in 1:length(specp$species)){set_para(i,shape=TRUE, setdf=specp)}
```

Now we can run the simulations:

```{r runSIm, warning=FALSE, message=FALSE}
if(force ==FALSE & file.exists('sim2.RDS')){
  message(Sys.time(), ': Found simulations...Using old simulaitons...')
  sim = readRDS('sim2.RDS')
}else{
message(Sys.time(),': Running simulations...')
if(runParallel == TRUE){
sim = foreach(i=1:length(setdf$L), .combine=rbind) %dopar% {set_para(i, shape=FALSE,setdf=setdf)}
}else{
sim = do.call('rbind',
pbapply(matrix(1:length(setdf$L)),1,function(i){set_para(i, shape=FALSE,setdf=setdf)})
)
}
message(Sys.time(),': Running simulations completed!')
saveRDS(sim, file='sim2.RDS', compress=TRUE)
}
```
  
Let's have a look at the results.

Let's plot TS grouped by species vs Length:
  
```{r fig.width=12, fig.height=12}
plot_func <- function(sim, name='L (mm)', var='L', y='TS',x='freq', title) {
  ggplot(data = sim, aes_string(x = x, y = y, col = var)) +
    geom_point(size=0.01) +
    scale_colour_viridis_c(name=name)+
    ggtitle(title)+
    xlab('Frequency (kHz)')+
    ylab('TS (dB)')+
    theme_classic()
}
pL <- function(sim,spec){plot_func(sim, name='L (mm)', var='L', y='TS',x='freq', spec)}

sim_tmp <- sim %>% 
  group_by(spec) %>% 
  nest() %>% 
  mutate(plots = map2(data, spec, pL))
pL = gridExtra::grid.arrange(grobs = sim_tmp$plots)
```


To have a look at the overall distribution, we can also look at 2D density plots:  

```{r fig.width=12, fig.height=12}
pL <- function(sim, title){

  ggplot(data=sim, aes(x=freq, y=TS))+
    geom_density2d_filled()+
    #facet_wrap(.~spec, scales='free')+
    scale_x_continuous(expand=c(0,0))+
    scale_y_continuous(expand=c(0,0))+
    ggtitle(title)+
    theme_classic()
}


sim_tmp <- sim%>% 
  group_by(spec) %>% 
  nest() %>% 
  mutate(plots = map2(data, spec, pL))
pL = gridExtra::grid.arrange(grobs = sim_tmp$plots)
```


```{r export-feather}
path <- "AZKABAN_ZoopMix_data_plusjelly.feather"
write_feather(sim, path)


```
